On the validity of the Born approximation for Weak Lensing observables

### 2016-11-02 Paper/paper.tex
switch order of smoothing and shape noise
routine for plotting skew(thetaG)
finished second version of the draft

\begin{abstract}
Accurate forward modeling of Weak Lensing (WL) observables from cosmological parameters is a necessary tool for upcoming galaxy surveys. Because WL probes structures in the non--linear regime, analytical forward modeling is very challenging, if not impossible. Numerical simulations of WL features rely on ray--tracing through the outputs of $N$--body simulations, which requires knowledge of the gravitational potential and accurate solvers for light ray trajectories. A cheaper but less accurate procedure based on the Born approximation, on the other hand, requires knowledge of the density only and can be implemented more efficiently and at lower computational cost. In this work we show that geodesic perturbations are the dominant correction to the Born approximation, lens-lens couplings being negligible. We also show that, although the Born approximation does not accurately predict WL convergence moments in the noiseless case, once galaxy shape noise is added the loss of accuracy is negligible even for a LSST--like survey, due to the increased statistical errors.    
\end{abstract}

@@

\begin{widetext}
\begin{equation}
\label{sim:jacsol}
A_{ij}(\pt,\chi_s) = \delta^K_{ij}-2\int_0^{\chi_s} d\chi\chi W(\chi,\chi_s)\Phi_{ik}(\bb{x}_\perp(\chi,\pt),\chi)A_{kj}(\pt,\chi)
\end{equation} 
\end{widetext}
%
We indicated the partial derivatives of the lensing potential $\Phi$ with respect to the transverse coordinates $x,y$ as subscripts. We also defined the single redshift source lensing kernel $W(\chi,\chi_s)=1-\chi/\chi_s$ and indicated the Kronecker delta symbol as $\delta^K$. The WL convergence $\kappa(\pt,\chi_s)$ can be calculated from the trace of the Jacobian
\begin{equation}
\label{sim:trkappa}
\kappa(\pt,\chi_s) = 1-\frac{1}{2}{\rm tr}\left[\bb{A}(\pt,\chi_s)\right]
\end{equation}
%
The implicit form of equation (\ref{sim:jacsol}) suggests a straightforward way to solve the geodesic equation numerically because $A(\pt,\chi_s)$ can be expressed once one knows $A(\pt,\chi)$ for $\chi<\chi_s$. The multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} is a popular method to compute the integral in (\ref{sim:jacsol}) in discrete steps (lenses) by keeping track of the intermediate values $A(\pt,\chi)$ using dynamic programming. To carry out the geodesic calculations we make use of the \ttt{LensTools} software package \citep{LensTools-paper}, which provides a {\sc python} implementation of the multi--lens--plane algorithm. The exact solution of (\ref{sim:geodiff}) based on (\ref{sim:jacsol}) is computationally expensive because it requires knowledge of the lensing potential $\Phi$. In the next paragraph we review the details of a computationally cheaper but approximate approach based on the Born approximation.

@@

\begin{equation}
\label{sim:born}
\kappa^{\rm born}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s} d\chi\frac{\chi}{a(\chi)} W(\chi,\chi_s)\delta(\chi\pt,\chi)
\end{equation}
%
The Born approximated convergence in equation (\ref{sim:born}) has a simple interpretation: at lowest order in the lensing potential, $\kappa$ is the integrated matter density on the unperturbed line of sight, weighted by the lensing kernel $W$. Contrary to the exact ray--tracing approach, the Born approximation does not required knowledge of the solution to the Poisson equation (\ref{sim:poisson}) and is hence computationally cheaper. In the next paragraph we examine the corrections to the Born approximation that come at second order in the potential $\Phi$.

@@

\begin{equation}
\label{sim:pb2}
\kappa = \kappa^{\rm born} + \kappa^{\rm ll} + \kappa^{\rm geo} + O(\Phi^3)
\end{equation}
%
The second order corrections can be expressed as double integrals along the unperturbed line of sight

@@

\begin{equation}
\label{sim:ll}
\kappa^{\rm ll}(\pt,\chi_s) = \int_0^{\chi_s}d\chi\int_0^\chi d\chi' \chi\chi' W(\chi,\chi_s)W(\chi',\chi_s)\Phi_{ij}(\chi\pt,\chi)\Phi_{ij}(\chi'\pt,\chi')
\end{equation}

\begin{equation}
\label{sim:gp}
\kappa^{\rm geo}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s}d\chi\int_0^\chi d\chi' \frac{\chi\chi'}{a(\chi')} W(\chi,\chi_s)W(\chi',\chi_s)\nabla_\perp\Phi(\chi\pt,\chi)\cdot\nabla_\perp\delta(\chi'\pt,\chi')
\end{equation}

\end{widetext}
%
These second order expression have a simple physical interpretation. Equation (\ref{sim:ll}) encodes the post--Born correction to the convergence due to non local quadratic lens--lens couplings, which add to the density line of sight integration in (\ref{sim:born}). Equation (\ref{sim:gp}), on the other hand, encodes the integral of the density contrast on the real light ray trajectory, at lowest order in the geodesic deflections. The line of sight integrals (\ref{sim:born}), (\ref{sim:ll}) and (\ref{sim:gp}) can be computed efficiently with runtime that scales linearly with the number of lenses using the functionality of the \ttt{LensTools} suite \citep{LensTools-paper}. A sample of the simulation outputs is shown in Figure \ref{fig:csample}. In the next section we show how to infer cosmological parameters from WL features forward models applied to observations.

@@

\section{Results}
% 
\begin{figure*}
\begin{center}
\includegraphics[scale=0.3]{Figures/powerResiduals.eps}
\end{center}
\caption{Contributions of different $\Phi$ orders to the $\kappa$ power spectrum: the left panel shows the auto power spectra $P_{\rm born,born}$ ($O(\Phi^2)$) and $P_{\rm ll},P_{\rm geo}$ ($O(\Phi^4)$). The right panel shows the residuals between the power spectrum of the full ray--traced $\kappa$ field and the one obtained with the Born approximation. The residuals are compared to the two $O(\Phi^3)$ terms $2P_{\rm born,ll},2P_{\rm born,geo}$. The quantities shown are the averages over 1024 realizations of $\kappa$ in the fiducial cosmology $\bb{p}_0$.}
\label{fig:psResiduals}
\end{figure*}
%
We also studied the accuracy of the Born approximation in predicting the 9 higher order convergence moments defined in (\ref{par:moments}). The results are shown in Figure \ref{fig:pdfMoments}.

@@

@@

In this section we presents the results of this work, regarding the WL feature forward model accuracy and the cosmological parameter bias induced by the Born approximation.

In this paragraph we compare the accuracy of the Born approximation in predicting WL image features, in particular we focus on the $\kappa$ power spectrum and the 9 moments described in equation (\ref{par:moments}). Figure \ref{fig:psResiduals} shows the auto power spectra of $\kappa^{\rm born},\kappa^{\rm ll},\kappa^{\rm geo}$ as defined in equation (\ref{sim:pb2}). At lowest order the lensing potential contributes to the $\kappa$ power spectrum quadratically. The first post--Born correction to $P^{\kappa\kappa}$ comes to $O(\Phi^3)$ and is given by
\begin{equation}
\label{res:powercorr}
P^{\kappa\kappa} = P^{\kappa\kappa}_{\rm born,born} + 2P^{\kappa\kappa}_{\rm born,ll} + 2P^{\kappa\kappa}_{\rm born,geo} + O(\Phi^4)
\end{equation}
%
Where we defined the $\kappa$ cross spectra as
\begin{equation}
\label{res:powercross}
\langle\tilde{\kappa}^{a}(\pmb{\ell})\tilde{\kappa}^{b,*}(\pmb{\ell}')\rangle = (2\pi)^2\delta^D(\pmb{\ell}+\pmb{\ell}')P_{a,b}^{\kappa\kappa}(\ell)
\end{equation} 
%
Figure \ref{fig:psResiduals} shows the magnitudes of the $O(\Phi^3)$ post--Born corrections to the convergence power spectrum.

@@

\section{Discussion}
% 
In this section we discuss our main findings. Figure \ref{fig:csample} shows that the dominant post--Born correction to the $\kappa$ reconstruction comes from the geodesic term (\ref{sim:gp}). This can be shown both looking at the lens-lens term, which is one order of magnitude smaller than the geodesic term, and looking at the $\kappa-\kappa^{\rm born}$ residuals, which show similar structure to the ones found in the $\kappa^{\rm geo}$ map.

Figure \ref{fig:psResiduals} shows a comparison between the Born approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. Figure \ref{fig:psResiduals} also gives indications that the perturbation series in powers of $\Phi$ becomes asymptotic on small angular scales as the higher orders $O(\Phi^4)$ tend to dominate the leading orders. This is likely due to the non--linearity of the density contrast on small scales.

Figure \ref{fig:pdfMoments} shows the PDF of the 9 $\kappa$ moments defined in (\ref{par:moments}) over 1000 realizations of an LSST--like survey. The plot clearly shows that, while the Born approximation is a good predictor of the quadratic moments $\pmb{\mu}_2$ within one standard deviation, the higher moments $\pmb{\mu}_3,\pmb{\mu}_4$ show $2\div 3\sigma$ deviations from their Born approximated counterparts. This can lead to parameter biases when fitting observations. Figure \ref{fig:parbias} studies this possibility and shows that, while the Born approximation is sufficient to fit cosmology with the power spectrum of an LSST--like survey, even in the noiseless case, the same is not true for the higher $\kappa$ moments. The middle panel of Figure \ref{fig:parbias} show that high significance biases are observed when fitting a real observation with Born approximated $\kappa$ moments. The bias however disappears when introducing galaxy shape noise, due to the resulting increased statistical error.

@@

\begin{itemize}
\item The main corrections to the Born approximations come from the perturbations in the light ray geodesic, with lens--lens coupling being at least one order of magnitude sub leading
\item The $\kappa$ power spectrum leads to negligible parameter bias for an LSST--like survey, even in the noiseless case
\item Fitting an observation with Born approximated higher $\kappa$ moments leads to high significance bias in the $(\Omega_m,w_0,\sigma_8)$ in the noiseless case. The bias significance becomes negligible when galaxy shape noise is added due to an increased scatter in parameter estimates.   
\end{itemize}
%
Future prospects involve study the Born approximation validity for CMB lensing observations.

@@

\section*{Acknowledgments}
We thank Uros Seljak for useful discussions. The $N$--body simulations in this work were performed at National Energy Research Scientific Computing Center (NERSC). We thank the LSST Dark Energy Science Collaboration (DESC) for the allocation of time. The ray--tracing simulations and parameter estimation calculations were performed at the NSF XSEDE facility, supported by grant number
ACI-1053575, and at the New York Center for Computational Sciences, a
cooperative effort between Brookhaven National Laboratory and Stony
Brook University, supported in part by the State of New York. This
work was supported in part by the U.S. Department of Energy under
Contract No. DE-SC00112704, and by the NSF Grant
No. AST-1210877 (to Z.H.) and by the Research Opportunities and
Approaches to Data Science (ROADS) program at the Institute for Data
Sciences and Engineering at Columbia University (to Z.H.).

>>>>

\begin{abstract}
Accurate forward modeling of Weak Lensing (WL) observables from cosmological parameters is a necessary tool for upcoming galaxy surveys. Because WL probes structures in the non--linear regime, analytical forward modeling is very challenging, if not impossible. Numerical simulations of WL features rely on ray--tracing through the outputs of $N$--body simulations, which requires knowledge of the gravitational potential and accurate solvers for light ray trajectories. A cheaper but less accurate procedure based on the Born approximation, on the other hand, requires knowledge of the density only and can be implemented more efficiently and at lower computational cost. In this work we show that the deviations of the Born--approximated convergence power spectrum, skewness and kurtosis from the ray--traced ones are consistent with the smallest non trivial post--Born corrections. We also show that perturbative approaches to the post--Born corrections break down at higher orders, making full ray--tracing more reliable. Because we find that cosmological parameter biases induced by the Born approximation are negligible even for an LSST--like survey once galaxy shape noise is considered, we argue that the Born approximation is preferred for future analysis. Using the \ttt{LensTools} software suite, we show that the Born approximation saves a factor of 4 in computing time with respect to the full ray--tracing in reconstructing the convergence.       
\end{abstract}

@@

\begin{widetext}
\begin{equation}
\label{sim:jacsol}
A_{ij}(\pt,\chi_s) = \delta^K_{ij}-2\int_0^{\chi_s} d\chi\chi W(\chi,\chi_s)\Phi_{ik}(\bb{x}_\perp(\chi,\pt),\chi)A_{kj}(\pt,\chi)
\end{equation} 
\end{widetext}
%
We indicated the partial derivatives of the lensing potential $\Phi$ with respect to the transverse coordinates $x,y$ as subscripts. We also defined the single redshift source lensing kernel $W(\chi,\chi_s)=1-\chi/\chi_s$ and indicated the Kronecker delta symbol as $\delta^K$. The WL convergence $\kappa(\pt,\chi_s)$ can be calculated from the trace of the Jacobian
\begin{equation}
\label{sim:trkappa}
\kappa(\pt,\chi_s) = 1-\frac{1}{2}{\rm tr}\left[\bb{A}(\pt,\chi_s)\right]
\end{equation}
%
The implicit form of equation (\ref{sim:jacsol}) suggests a straightforward way to solve the geodesic equation numerically because $A(\pt,\chi_s)$ can be calculated once $A(\pt,\chi)$ for $\chi<\chi_s$ is known. The multi--lens--plane algorithm \citep{RayTracingJain,RayTracingHartlap} is a popular method to compute the integral in (\ref{sim:jacsol}) in discrete steps (lenses) by keeping track of the intermediate values $A(\pt,\chi)$ using dynamic programming. To carry out the geodesic calculations we make use of the \ttt{LensTools} software package \citep{LensTools-paper}, which provides a {\sc python} implementation of the multi--lens--plane algorithm. The exact solution of (\ref{sim:geodiff}) based on (\ref{sim:jacsol}) is computationally expensive because it requires knowledge of the lensing potential $\Phi$. In the next paragraph we review the details of a computationally cheaper but approximate approach based on the Born approximation.

@@

\begin{equation}
\label{sim:born}
\kappa_{\rm born}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s} d\chi\frac{\chi}{a(\chi)} W(\chi,\chi_s)\delta(\chi\pt,\chi)
\end{equation}
%
The Born approximated convergence in equation (\ref{sim:born}) has a simple interpretation: at lowest order in the lensing potential, $\kappa$ is the integrated matter density on the unperturbed line of sight, weighted by the lensing kernel $W$. Contrary to the exact ray--tracing approach, the Born approximation does not required knowledge of the solution to the Poisson equation (\ref{sim:poisson}) and is hence computationally cheaper. In the next paragraph we examine the corrections to the Born approximation that come at second order in the potential $\Phi$.

@@

\begin{equation}
\label{sim:pb2}
\kappa = \kappa_{\rm born} + \kappa_{\rm ll} + \kappa_{\rm geo} + O(\Phi^3)
\end{equation}
%
The second order corrections can be expressed as double integrals along the unperturbed line of sight

@@

\begin{equation}
\label{sim:ll}
\kappa_{\rm ll}(\pt,\chi_s) = \int_0^{\chi_s}d\chi\int_0^\chi d\chi' \chi\chi' W(\chi,\chi_s)W(\chi',\chi)\Phi_{ij}(\chi\pt,\chi)\Phi_{ij}(\chi'\pt,\chi')
\end{equation}

\begin{equation}
\label{sim:gp}
\kappa_{\rm geo}(\pt,\chi_s) = \frac{3H_0^2\Omega_m}{2c^2}\int_0^{\chi_s}d\chi\int_0^\chi d\chi' \frac{\chi\chi'}{a(\chi')} W(\chi,\chi_s)W(\chi',\chi)\nabla_\perp\Phi(\chi\pt,\chi)\cdot\nabla_\perp\delta(\chi'\pt,\chi')
\end{equation}

\end{widetext}
%
These second order expression have a simple physical interpretation. Equation (\ref{sim:ll}) encodes the post--Born correction to the convergence due to non local quadratic lens--lens couplings, which add to the density line of sight integration in (\ref{sim:born}). Equation (\ref{sim:gp}), on the other hand, encodes the integral of the density contrast on the real light ray trajectory, at lowest order in the geodesic deflections. The line of sight integrals (\ref{sim:born}), (\ref{sim:ll}) and (\ref{sim:gp}) can be computed efficiently with runtime that scales linearly with the number of lenses using the functionality of the \ttt{LensTools} suite \citep{LensTools-paper}. A sample of the simulation outputs is shown in Figure \ref{fig:csample}. Benchmarks for the runtime of the ray--tracing and line--of--sight integrals operations, performed with \ttt{LensTools}, are provided in Table \ref{tab:benchmarks}. 
In the next section we show how to infer cosmological parameters from WL features forward models applied to observations.

\begin{table*}
\begin{center}

\textbf{Integration type} & \textbf{Runtime (One field of view)} & \textbf{Memory usage} & \textbf{CPU time (1000 fields of view)} \\ \hline \hline
Born & 36.0\,s & 0.86\,GB & 10\,hours  \\
Full ray--tracing & 124.8\,s & 1.65\,GB & 35\,hours  \\
Born + $O(\Phi^2)$ & 156.7\,s & 1.52\,GB & 44\,hours \\ \hline

\caption{Time and memory usage benchmarks for the convergence reconstruction operations. The test case we considered consists in a discretization with 42 uniformly spaced lenses between the observer and the sources at $z_s=2$, each with a resolution of $4096^2$ pixels. The $\kappa$ field is resolved with $2048^2$ light rays. We show both the runtime for producing a single field of view and the CPU hours needed to perform the reconstruction 1000 times, to mock an LSST--like galaxy survey.}
\label{tab:benchmarks}

\end{table*}

@@

\section{Results}
% 
In this section we presents the results of this work, regarding the WL feature forward model accuracy and the cosmological parameter bias induced by the Born approximation.

\begin{figure*}
\begin{center}
\includegraphics[scale=0.3]{Figures/powerResiduals.eps}
\end{center}
\caption{Contributions of different $\Phi$ orders to the $\kappa$ power spectrum: the left panel shows the auto power spectra $P_{\rm born,born}$ ($O(\Phi^2)$) and $P_{\rm ll},P_{\rm geo}$ ($O(\Phi^4)$). The right panel shows the residuals between the power spectrum of the full ray--traced $\kappa$ field and the one obtained with the Born approximation. The residuals are compared to the two $O(\Phi^3)$ terms $2P_{\rm born,ll},2P_{\rm born,geo}$. The quantities shown are the ensemble averages over 8192 realizations of $\kappa$ in the fiducial cosmology $\bb{p}_0$.}
\label{fig:psResiduals}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[scale=0.4]{Figures/delta_m2.eps}
\includegraphics[scale=0.4]{Figures/delta_m5.eps}
\end{center}
\caption{Dominant post--Born corrections to the $\kappa$ skewness and kurtosis due to lens--lens coupling and geodesic perturbation, as a function of the scale $\theta_G$ of the Gaussian smoothing window. We show the contributions of the lens--lens couplings (red) and the geodesic perturbation (green). We show both the cases in which the residuals $\delta\kappa^3,\delta\kappa^4$ are computed adding equations (\ref{sim:ll},\ref{sim:gp}) to the Born field (solid lines) and in which the residuals are computed from the dominant $\Phi$ orders only as in equations (\ref{res:skewcross},\ref{res:kurtcross}) (dashed lines). The quantities shown are the ensemble averages over 8192 realizations of $\kappa$ in the fiducial cosmology $\bb{p}_0$.}
\label{fig:skResiduals}
\end{figure*}

@@

In this paragraph we compare the accuracy of the Born approximation in predicting WL image features, in particular we focus on the $\kappa$ power spectrum and the 9 moments described in equation (\ref{par:moments}). Figure \ref{fig:psResiduals} shows the auto power spectra of $\kappa_{\rm born},\kappa_{\rm ll},\kappa_{\rm geo}$ as defined in equation (\ref{sim:pb2}). At lowest order the lensing potential contributes to the $\kappa$ power spectrum quadratically. The first post--Born correction to $P^{\kappa\kappa}$ comes to $O(\Phi^3)$ and is given by
\begin{equation}
\label{res:powercorr}
P^{\kappa\kappa} = P^{\kappa\kappa}_{\rm born,born} + 2P^{\kappa\kappa}_{\rm born,ll} + 2P^{\kappa\kappa}_{\rm born,geo} + O(\Phi^4)
\end{equation}
%
Where we defined the $\kappa$ cross spectra as
\begin{equation}
\label{res:powercross}
\langle\tilde{\kappa}_a(\pmb{\ell})\tilde{\kappa}_b^*(\pmb{\ell}')\rangle = (2\pi)^2\delta^D(\pmb{\ell}+\pmb{\ell}')P_{a,b}^{\kappa\kappa}(\ell)
\end{equation} 
%
Figure \ref{fig:psResiduals} shows the magnitudes of the $O(\Phi^3)$ post--Born corrections to the convergence power spectrum. 
A similar analysis can be carried out for the skewness and kurtosis of the $\kappa$ field. The main contributions to the skewness and kurtosis are $O(\Phi^3),O(\Phi^4)$ respectively. The first post--Born corrections, on the other hand, come at $O(\Phi^5),O(\Phi^6)$. More in detail we have

\begin{equation}
\label{res:skewcross}
\langle\kappa^3\rangle = \langle\kappa_{\rm born}^3\rangle + 3\langle\kappa_{\rm born}^2\kappa_{\rm ll}\rangle + 3\langle\kappa_{\rm born}^2\kappa_{\rm geo}\rangle + O(\Phi^5) 
\end{equation}

\begin{equation}
\label{res:kurtcross}
\langle\kappa^4\rangle_c = \langle\kappa_{\rm born}^4\rangle_c + 4\langle\kappa_{\rm born}^3\kappa_{\rm ll}\rangle_c + 4\langle\kappa_{\rm born}^3\kappa_{\rm geo}\rangle_c + O(\Phi^6)
\end{equation}
%
Where the kurtosis connected components are defined as 
\begin{equation}
\label{res:kurtconnected}
\langle\kappa^3_a\kappa_b\rangle_c = \langle\kappa^3_a\kappa_b\rangle - 3\langle\kappa_a^2\rangle\langle\kappa_a\kappa_b\rangle
\end{equation}
The magnitude of the post--Born corrections to the $\kappa$ skewness and kurtosis in equations (\ref{res:skewcross},\ref{res:kurtcross}) are shown in Figure \ref{fig:skResiduals}.
We also studied the accuracy of the Born approximation in predicting the 9 higher order convergence moments defined in (\ref{par:moments}). The results are shown in Figure \ref{fig:pdfMoments}.

@@

@@

\section{Discussion}
% 
In this section we discuss our main findings. Figure \ref{fig:csample} shows that the dominant post--Born correction to the $\kappa$ reconstruction comes from the geodesic term (\ref{sim:gp}). This can be shown both looking at the lens-lens term, which is one order of magnitude smaller than the geodesic term, and looking at the $\kappa-\kappa_{\rm born}$ residuals, which show similar structure to the ones found in the $\kappa_{\rm geo}$ map.

Figure \ref{fig:psResiduals} shows a comparison between the Born approximated $\kappa$ power spectra and the ones obtained with full ray--tracing, separating different orders in $\Phi$. We can conclude that the dominant contributions to the residuals come from the Born--geodesic cross power, which dominates the Born--lens--lens cross power by two orders of magnitude on small scales. Figure \ref{fig:skResiduals} shows that the geodesic and lens--lens contributions to the $\kappa$ skewness and kurtosis are comparable and that the lowest non trivial post--Born corrections can fully account for the discrepancy between the Born approximated quantities and the ray--traced ones.  Both Figures \ref{fig:psResiduals} and \ref{fig:skResiduals} give indications about the fact that the post--Born perturbation series break down at high orders in the lensing potential. In particular geodesic perturbations cannot be predicted accurately by a perturbation series, higher orders in $\Phi$ being dominant over smaller orders. We do not see this as an important limitation as the computational cost of the $O(\Phi^2)$ line--of--sight integrals is roughly the same as the ray--tracing one, as shown in Table \ref{tab:benchmarks}.

Figure \ref{fig:pdfMoments} shows the PDF of the 9 $\kappa$ moments defined in (\ref{par:moments}) over 1000 realizations of an LSST--like survey. The plot clearly shows that, while the Born approximation is a good predictor of the quadratic moments $\pmb{\mu}_2$ within one standard deviation, the higher moments $\pmb{\mu}_3,\pmb{\mu}_4$ show $2\div 3\sigma$ deviations from their Born approximated counterparts. This can lead to parameter biases when fitting observations. Figure \ref{fig:parbias} studies this possibility and shows that, while the Born approximation is sufficient to fit cosmology with the power spectrum of an LSST--like survey, even in the noiseless case, the same is not true for the higher $\kappa$ moments. The middle panel of Figure \ref{fig:parbias} shows that high significance biases are observed when fitting a real observation with Born approximated $\kappa$ moments. The bias however disappears when introducing galaxy shape noise, due to the resulting increased statistical error. This result leads us to the conclusion that the Born approximation is sufficient for future WL analyses. This approach also has the advantage of being 4 times computationally cheaper than full ray--tracing, as we show in Table \ref{tab:benchmarks} for a selected reference test case.

@@

\begin{itemize}
\item The lowest order post--Born corrections to the convergence power spectrum, skewness and kurtosis are sufficient to fully account for the discrepancies between ray--tracing and the Born approximation
\item Using the \ttt{LensTools} software package, Born integration costs 4 times less than ray--tracing 
\item The $\kappa$ power spectrum leads to negligible parameter bias for an LSST--like survey, even in the noiseless case
\item Fitting an observation with Born approximated higher $\kappa$ moments leads to high significance bias in the $(\Omega_m,w_0,\sigma_8)$ in the noiseless case. The bias significance becomes negligible when galaxy shape noise is added due to an increased scatter in parameter estimates.   
\end{itemize}
%
Future prospects involve study the Born approximation validity for CMB lensing observations.

@@

\section*{Acknowledgments}
We thank Uros Seljak for useful discussions. The $N$--body simulations in this work were performed at National Energy Research Scientific Computing Center (NERSC). We thank the LSST Dark Energy Science Collaboration (DESC) for the allocation of time. The ray--tracing simulations and parameter estimation calculations were performed at the NSF XSEDE facility, supported by grant number
ACI-1053575, and at the New York Center for Computational Sciences, a
cooperative effort between Brookhaven National Laboratory and Stony
Brook University, supported in part by the State of New York. This
work was supported in part by the U.S. Department of Energy under
Contract No. DE-SC00112704, the NSF Grant
No. AST-1210877 (to Z.H.), the Research Opportunities and
Approaches to Data Science (ROADS) program at the Institute for Data
Sciences and Engineering at Columbia University (to Z.H.) and a Simons Fellowship (to Z.H.).
